# Introduction to Exploratory Cognitive Diagnostic Models {#intro}

Exploratory Cognitive Diagnostic Models (ECDMs) are versions of classicial 
Cognitive Diagnostic Models (CDMs) that do _not_ require the component of 
an expertly crafted _Q_ matrix. This class of models is new to the world of
psychometric models. The goal of this textbook is to provide an overview
of their implementation in the `ecdm` package from by the authors that 
developed it!

Before we continue, please "bookmark" the `ecdm` repository on GitHub:

<https://github.com/tmsalab/ecdm>

The website provides direct access to the developers behind `ecdm`. In particular,
it features the ability to file issues or bug reports, ask questions, or stay
up-to-date in the latest breakthroughs.

## Installation

Before we can get started, please install the `ecdm` package from GitHub.
The `ecdm` package is currently only available via GitHub as
it is still being developed. As a result, installing by
`install.packages('ecdm')` isn't possible. 

As many of the routines are written in _C++_, the `ecdm` package
requires a compiler to install. To assist in setting up the compiler, we've
created the following guides:

- [Windows: Rtools](http://thecoatlessprofessor.com/programming/installing-rtools-for-compiled-code-via-rcpp/)
- [macOS: Rtools](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-macos/)

From there, please use the `remotes` package to retrieve the latest development version.

```{r gh-installation, eval = FALSE}
if(!requireNamespace('remotes', quietly = TRUE)) install.packages("remotes")
remotes::install_github("tmsalab/ecdm")
```

## Loading the Package

Accessing the `ecdm` rountines requires loading the package into _R_. Please
load the `ecdm` package by pressing "run"

```{r main-lib}
library(ecdm)
```
