# Introduction to Exploratory Cognitive Diagnostic Models {#intro}

Exploratory Cognitive Diagnostic Models (ECDMs) are versions of classicial 
Cognitive Diagnostic Models (CDMs) that do _not_ require the component of 
an expertly crafted _Q_ matrix. This class of models is new to the world of
psychometric models. The goal of this textbook is to provide an overview
of their implementation in the `ecdm` package from by the authors that 
developed it!

Before we continue, please "bookmark" the `ecdm` repository on GitHub:

<https://github.com/tmsalab/ecdm>

The website provides direct access to the developers behind `ecdm`. In particular,
it features the ability to file issues or bug reports, ask questions, or stay
up-to-date in the latest breakthroughs.

## Installation

Before we can get started, please install the `ecdm` package from GitHub.
The `ecdm` package is currently only available via GitHub as
it is still being developed. As a result, installing by
`install.packages('ecdm')` isn't possible. 

As many of the routines are written in _C++_, the `ecdm` package
requires a compiler to install. To assist in setting up the compiler, we've
created the following guides:

- [Windows: Rtools](http://thecoatlessprofessor.com/programming/installing-rtools-for-compiled-code-via-rcpp/)
- [macOS: Rtools](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-macos/)

From there, please use the `remotes` package to retrieve the latest development version.

```{r gh-installation, eval = FALSE}
if(!requireNamespace('remotes', quietly = TRUE)) install.packages("remotes")
remotes::install_github("tmsalab/ecdm")
```

## Loading the Package

Accessing the `ecdm` rountines requires loading the package into _R_. Please
load the `ecdm` package by pressing "run"

```{r main-lib}
library(ecdm)
```

## Supplementary Data Sets

The `ecdm` package has an accompanying data package called `ecdmdata` that
comes equipped with three different data sets:

- Examination for the Certificate of Proficiency in English (ECPE), Templin, J. and Hoffman, L. (2013). 
    - `items_ecpe`, N = 2922 subject responses to J = 28 items.
    - `qmatrix_ecpe`, J = 28 items and K = 3 traits. 
- Fraction Addition and Subtraction, Tatsuoka, K. K. (1984).
    - `items_fractions`: N = 536 subject responses to J = 20 items.
    - `qmatrix_fractions`: J = 20 items and K = 8 traits. 
- Revised PSVT:R, Culpepper and Balamuta (2013).
    - `items_spatial`: N = 516 subject responses to J = 30 items.
    
As time goes on, more data sets will likely be added.

Let's take a look at Fraction Addition and Subtraction data sets. Typing the
name of each data set and running the command will load the data into _R_ if
the `ecdmdata` package is loaded. As these data sets are relatively big, 
let's use the function `head()` to view on the first 6 rows.

```{r data-inc}
library(ecdmdata)

head( items_fractions )
head( qmatrix_fractions )
```

Within this textbook, the following notation will be used for dimensions:

- $K$: Number of Traits (columns) in the Q Matrix
- $J$: Number of Items (rows/columns) in the Q Matrix and Item Matrix
- $N$: Number of Subjects (rows) in the Item Matrix

To retrieve this information in _R_, we can use the dimension function,
`dim()`, which lists the size of the data as rows by columns. 

Find the dimensions of the `items_fractions` and `qmatrix_fractions`

```{r dim-info}
dim(items_fractions)
dim(qmatrix_fractions)
```

## Help!

Each function within the package contains a help file that provides documentation
on the implementation. Moreover, some of the functions have _worked_ examples 
as well. To view this information type either `?function_name` or 
`help(function_name)`. Let's verify the previously acquired numbers for the 
`items_fractions` data set by checking the entry in the documentation.

```{r help-info}
?items_fractions
```

If you are curious to see how a function performs, you can opt to use 
`example(function_name, package = "ecdm")`. Be aware that some examples
may take considerably longer than the rest to run. 

## Notation

For consistency, we aim to use the following notation.

Denoting individuals:

- $N$ is the total number of individuals taking the assessment.
- $i$ is the current individual.

Denoting items:

- $J$ is the total number of items on the assessment.
- $j$ is the current item
- $Y_{ij}$ is the observed binary response for individual $i$ ($1\leq i \leq N$) to item $j$ ($1\leq j\leq J$).
- $s_j$ is the probability of slipping on item $j$. 
- $g_j$ is the probability of guessing on item $j$.

Denoting attributes:

- $K$ is the total number of attributes for the assessment item.
- $k$ is the current attribute.
- $\boldsymbol\alpha_i=\left(\alpha_{i1},\dots,\alpha_{iK}\right)^\prime$ 
  where $\boldsymbol\alpha_i\in \left\{0,1\right\}^K$ and $\alpha_{ik}$ is 
  the latent binary attribute for individual $i$ on attribute $k$ ($1\leq k\leq K$).

Denoting the skill/attribute "Q" matrix: 

- $\boldsymbol q_{j}=\left(q_{j1},\dots,q_{jK}\right)^\prime$ be the
  $j$th row of $\boldsymbol Q$ such that $q_{jk}=1$ if 
  attribute $k$ is required for item $j$ and zero otherwise.

